V1298~Tau with TESS: Updated Ephemerides, Radii, and Period Constraints for a Previous Single Transit Event of V1298~Tau~e

### 2021-10-23 src/ms.tex
static, since no data right now
rearranging & new figure script

\title{V1298 Tau with TESS: Updated Ephemerides, Radii, and Period Constraints for V1298~Tau~e}

@@

\author{Charles Beichman}
\affiliation{Caltech/IPAC, 1200 E. California Blvd. Pasadena, CA 91125, USA}

\author[0000-0002-3199-2888]{Sarah Blunt}
\altaffiliation{NSF Graduate Research Fellow}
\affiliation{Department of Astronomy, California Institute of Technology, Pasadena, CA, USA}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.46\textwidth,trim={0.25cm 0 0 0}]{figures/TESSaperture.pdf}
\caption{The \tess\ target pixel file (TPF) overlaid with a sky image of \sname taken with the Sloan Digital Sky Survey (SDSS) i-band. \sname is shown in the white circle; nearby sources with \tess\ magnitudes $< 14$ are marked with orange x's. There are two bight nearby sources to \sname. However, selecting the one pixel aperture (white square) encapsulates only \sname and therefore contamination from the nearby sources is negligible. \href{https://github.com/afeinstein20/v1298tau_tess/blob/main/src/figures/tpf.py}{\github}} \label{fig:tpf}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht]
\begin{center}
\includegraphics[width=\textwidth,trim={0.25cm 0 0 0}]{figures/transits.pdf}
\caption{\sname extracted light curve from the \texttt{tica}-processed full-frame images, with transits of \allplanets highlighted by color. Top row: extracted light curve with over plotted with our best-fit GP model for stellar variability (blue). Middle row: the \tess\ light curve with the stellar variability removed by our model. Bottom rows: zoomed-in regions around the visible transits during the first orbit (third row) and second orbit (last row) from \tess\ Sector 43.} \label{fig:transits}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

where $y$ is the flux array, and $\sigma_l$  and $\sigma_j$ are used to define the light curve noise and in-transit jitter. $\sigma_l$  and $\sigma_j$ are also used as the first and second terms in our rotation model, which we defined as a stochastically-driven, damped harmonic oscillator, defined by the \texttt{SHOTerm} in \texttt{celerite2} \citep{dfm17}.

@@

There was a single transit of \planete in the original \textit{K2} data, which occurred roughly in the middle of the campaign. Since no other transits were detected, this provides a lower period limit of 36~days. Using the original transit timing from \textit{K2} and this new transit timing from \tess, we place further constraints on the period by sampling over discrete periods, similarly to \cite{becker19}. For this analysis, we used the \texttt{EVEREST 2.0} \citep{luger18} version of the \textit{K2} light curve.

We localized a 1-day window around each transit of \planete and ran a GP fit using \texttt{exoplanet} for the transit parameters, assuming a constant depth between the \textit{K2} and \tess\ observations. We assumed there is no correlation between transit parameters fit and the period we are fitting for. We then compute possible periods 
\begin{equation}\label{eq:period}
    P = \frac{1}{q} \left(T_{mid,TESS} - T_{mid, K2}\right)
\end{equation}
where $T_{mid}$ are the transit midpoints from \textit{K2} and \tess\ and $q$ is an integer. We assume a uniform prior, i.e. we have no prior preference for a specific harmonic, $q$. At each step of the sampling process, we compute a new light curve with different orbital periods, given by Equation~\ref{eq:period}. The log likelihood of each individual light curve, i.e. period, is saved, along with the sum of all log likelihoods for each period. The summation of all log likelihoods is used to compute the posterior probability for each sampled value of $q$.

We chose to test periods from $40-56$~days. We ran 3 chains with 500 tuning steps and 3000 draws. Similarly to the transit fits, we discarded the tuning steps before our analysis. We show our results in Figure~\ref{fig:period_e}, where we plot the median period for each tested harmonic against the posterior probability of each harmonic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth,trim={0.25cm 0 0cm 0}]{figures/periode.pdf}
\caption{Our calculated log likelihood to constrain the period of \planete using transit timings from \textit{K2} and \tess. We tested discrete periods from $44-56$\,days and find the most likely period to be \textcolor{red}{n}.} \label{fig:period_e}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We constrain the period of \planete to be \textcolor{red}{n}~days. Even at the lower limit, it is unlikely a transit of \planete will occur in the next sector (Sector 44) of \tess\ data for \sname. Independent ground-based monitoring of the system for several days beyond the end of Sector 44 (UT 06 Nov 2021) may be able to observe another transit of the outermost known planet in this system.

@@

If \allplanets host extended haze-dominated atmospheres, they should have smaller transit radii at mid-IR wavelengths \citep{gao20}. However, without mass estimates for \allplanets, it is difficult to determine if this is the cause of the different transit depths we measured between the \textit{K2} and \tess\ data. While radial velocity mass measurements are challenging for young planets due to the underlying stellar activity, efforts should be taken to try and constrain this parameter \citep{crockett12}.

@@

\software{\texttt{exoplanet} \citep{exoplanet2021},
          \texttt{EVEREST 2.0} \citep{luger18},
          \texttt{lightkurve} \citep{lightkurve},
          \texttt{matplotlib} \citep{matplotlib},
          \texttt{PyMC3} \citep{Salvatier16},
          \texttt{starry} \citep{luger19},
          \texttt{theano} \citep{theano},
          \texttt{tica} \citep{fausnaugh20},
          \texttt{batman} \citep{Kreidberg15},
          \texttt{showyourwork!} \citep{luger21},
          \texttt{astropy} \citep{astropy:2013, astropy18},
          \texttt{astroquery}\citep{astroquery19},
          numpy \citep{numpy}
          }

>>>>

\title{V1298~Tau with TESS: Updated Ephemerides, Radii, and Period Constraints for a Previous Single Transit Event of V1298~Tau~e}

@@

%\author{Charles Beichman}
%\affiliation{Caltech/IPAC, 1200 E. California Blvd. Pasadena, CA 91125, USA}

%\author[0000-0002-3199-2888]{Sarah Blunt}
%\altaffiliation{NSF Graduate Research Fellow}
%\affiliation{Department of Astronomy, California Institute of Technology, Pasadena, CA, USA}

@@

%An analysis of the Gaia DR1 astrometric data found that \sname was co-moving with 8 other stars \citep{Oh2017}. \citet{Luhman2018} used Gaia DR2 astrometry to expand the membership of this group, referred to as Oh's Group 29, and examine its relationship to the Taurus-Auriga star-forming region. \citet{Luhman2018} concluded that while Group 29 is indeed young ($\sim$40~Myr), the collection of stars was physically unrelated to the traditional star-forming groups within the Taurus-Auriga complex.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.46\textwidth,trim={0.25cm 0 0 0}]{static/TESSaperture.pdf}
\caption{The \tess\ target pixel file (TPF) overlaid with a sky image of \sname taken with the Sloan Digital Sky Survey (SDSS) i-band. \sname is shown in the white circle; nearby sources with \tess\ magnitudes $< 14$ are marked with orange x's. There are two bight nearby sources to \sname. However, selecting the one pixel aperture (white square) encapsulates only \sname and therefore contamination from the nearby sources is negligible. \href{https://github.com/afeinstein20/v1298tau_tess/blob/main/src/figures/tpf.py}{\github}} \label{fig:tpf}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[hbtp]
\begin{center}
\includegraphics[width=\textwidth,trim={0.25cm 0 0 0}]{figures/transits.pdf}
\caption{\sname extracted light curve from the \texttt{tica}-processed full-frame images, with transits of \allplanets highlighted by color. Top row: extracted light curve with over plotted with our best-fit GP model for stellar variability (blue). Middle row: the \tess\ light curve with the stellar variability removed by our model. Bottom rows: zoomed-in regions around the visible transits during the first orbit (third row) and second orbit (last row) from \tess\ Sector 43.} \label{fig:transits}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

where $y$ is the flux array normalized about zero, and $\sigma_l$  and $\sigma_j$ are used to define the light curve noise and in-transit jitter. $\sigma_l$  and $\sigma_j$ are also used as the first and second terms in our rotation model, which we defined as a stochastically-driven, damped harmonic oscillator, defined by the \texttt{SHOTerm} in \texttt{celerite2} \citep{dfm17}.

@@

There was a single transit of \planete in the original \textit{K2} data, which occurred roughly in the middle of the campaign. Since no other transits were detected, this provides a lower period limit of 36~days. Using the original transit timing from \textit{K2} and this new transit timing from \tess, we developed a new method for constraining the period of \planete. For this analysis, we used the \texttt{EVEREST 2.0} \citep{luger18} version of the \textit{K2} light curve.

For this analysis, we created a GP model that constrains the period of \planete by sampling over discrete periods, ranging from $38 - 56$~days. We fit for the additional transit parameters, $\theta$, assuming a constant transit depth between the \textit{K2} and \tess\ observations, $X$. To speed up the computation of the GP fits, we localized a 1-day window around each transit of \planete.

We assumed there is no correlation between the other transit parameters and the period we are fitting for. For each step in our MCMC fit, we compute all possible periods

\begin{equation}\label{eq:period}
    P = \frac{1}{q} \left(T_{mid,TESS} - T_{mid, K2}\right)
\end{equation}

where $T_{mid}$ are the transit midpoints from \textit{K2} and \tess\ and $q$ is an integer. We assume a uniform prior, i.e. we have no prior preference for a specific harmonic, $q$. At each step of the sampling process, we compute a new light curve with different orbital periods, given by Equation~\ref{eq:period}. The log likelihood of each individual generated light curve is saved

\begin{equation}
    \textrm{log} \mathcal{L}_q = \left[ \textrm{log}\, p \left( X | \theta^k, q^k = n \right) \right]_n
\end{equation}

along with the sum of all log likelihoods for each period

\begin{equation}
    \textrm{log} \mathcal{L} = \textrm{log}\, \Sigma_q\, p(q)\, p(X|\theta^k, q) 
\end{equation}

The summation of all log likelihoods is used to compute the posterior probability for each sampled value of $q$.

We chose to test periods from $38-56$~days. We ran 5 chains with 500 tuning steps and 5000 draws. Similarly to the transit fits, we discarded the tuning steps before our analysis. We show our results in Figure~\ref{fig:period_e}, where we plot the median period for each tested harmonic against the posterior probability of each harmonic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth,trim={0.25cm 0 0cm 0}]{figures/periode.pdf}
\caption{Our calculated log likelihood to constrain the period of \planete using transit timings from \textit{K2} and \tess. We tested discrete periods from $38-56$\,days and find the most likely period to be 50.748~days. The 2:1 resonance (48.28~days) with \planetb is plotted as the dashed vertical line.} \label{fig:period_e}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{deluxetable}{l r r}[!ht]
\tabletypesize{\footnotesize}
\tablecaption{Predicted Ground-based Transit Midpoint Events for \planete from Maunea Kea \label{tab:planete}}
\tablehead{
\textit{P} [days] & Posterior Prob. & Transit Dates [UT] \\
\hline
50.748 & $1.88 \times 10^{-4}$ &  \\
49.691 & $1.88 \times 10^{-4}$ & \\
48.677 & $1.15 \times 10^{-4}$ & \\
51.852 & $1.10 \times 10^{-4}$ &  \\
47.703 & $4.45 \times 10^{-5}$ & 
}
\startdata
\enddata
\tablecomments{Transit dates were calculated using the Transit Service Query Form on the NASA Exoplanet Archive.}
\end{deluxetable}

Using this method, we find the most likely period of \planete to be 50.748~days. We computed a $\chi^2$-fit to a Gaussian distribution and find the period to be $50.29 \pm 6.62$~days. Even at the lower limit, it is unlikely a transit of \planete will occur in the next sector (Sector 44) of \tess\ data for \sname. However, this derived period estimate suggests that \planete is in a near 2:1 mean motion resonance with \planetb.

Independent ground-based monitoring of the system for several days beyond the end of Sector 44 (UT 06 Nov 2021) may be able to observe another transit of the outermost known planet in this system. We provide several transit dates and times that are observable from Maunea Kea, given the top 5 most-likely periods of \planete, in Table~\ref{tab:planete}

@@

If \allplanets host extended haze-dominated atmospheres, they should have smaller transit radii at mid-IR wavelengths \citep{gao20}. However, without mass estimates for \allplanets, it is difficult to determine if this is the cause of the different transit depths we measured between the \textit{K2} and \tess\ data. While radial velocity mass measurements are challenging for young planets due to the underlying stellar activity, efforts should be taken to try and constrain the planet masses.

@@

\software{\texttt{exoplanet} \citep{exoplanet2021},
          \texttt{EVEREST 2.0} \citep{luger18},
          \texttt{lightkurve} \citep{lightkurve},
          \texttt{matplotlib} \citep{matplotlib},
          \texttt{PyMC3} \citep{Salvatier16},
          \texttt{starry} \citep{luger19},
          \texttt{theano} \citep{theano},
          \texttt{tica} \citep{fausnaugh20},
          \texttt{batman} \citep{Kreidberg15},
          \texttt{showyourwork!} \citep{luger21},
          \texttt{astropy} \citep{astropy:2013, astropy18},
          \texttt{astroquery}\citep{astroquery19},
          \texttt{numpy} \citep{numpy}
          }
